---
title: "Final Data Project"
author: "Alanna Bachtlin"
date: "7/29/2021"
output: html_document
header-includes:
     \usepackage{amsmath}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r include=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(latex2exp)
library(gridExtra)
library(grid)
```

   Data was collected using an acoustic Doppler velocimeter and a dissolved oxygen logger. For each site, one along the Allegheny River and two along the Monongahela River, graphs of height vs dissolved oxygen and height vs velocity are shown. The Reynolds Stress Tensor graphs are also shown for four different depths. From this data collection, the boundary layer can be studied. 
     
Reynolds Stress Tensor Equation:

\begin{equation}
\sigma_{ij} = \rho_0 \begin{bmatrix} \overline{u^2} & \overline{uv} & \overline{uw}\\ \overline{vu} & \overline{v^2} & \overline{vw}\\ \overline{wu} & \overline{wv} & \overline{w^2} \end{bmatrix}
\end{equation}

Normal stresses are diagonal in the tensor and sheer stresses are off-diagonal. The turbulent fluctuations exert a stress on the flow. If the graph appears isotrophic, there is no directional preference and the average mean of the fluctuations is zero but if the graph appears anisotrophic there is a directional preference and the mean fluctuations are not zero. 




```{r warning=FALSE, echo=FALSE}
#Allegheny River

#height vs DO
DO528AR214 <- read.csv("C:\\Users\\duquesne\\Documents\\Alanna\\DO_528AR214.csv")
DO528AR214 <- rename(DO528AR214, time = ï..time) # This appears to be needed for Windows... 

p1 <- ggplot(DO528AR214, aes(x = DO.mg.l. , y= height.m.))+
  geom_point() +
  labs(x = "DO (mg/l)", y = "Height (m)") +
  ggtitle("Dissolved Oxygen") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(axis.text = element_text(face = "plain", size = 8)) 
  
```

```{r warning=FALSE, echo=FALSE}
#height vs u
p2 <- ggplot(DO528AR214, aes(x = U.m.s. , y= height.m.))+
  geom_point() +
  labs(x = "Average Velocity (m/s)", y = "Height (m)") +
  ggtitle("Velocity") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(axis.text = element_text(face = "plain", size = 8)) 

```

```{r warning=FALSE, echo=FALSE}
#stress tensors 
#Allegheny River goes from 16:57:00- 17:21:00 in 2 minute increments
setwd("c:/Users/duquesne/Documents/nortek/data")
fh <- "528AR214" # filename header
fn_sen <- paste(fh, "sen", sep = ".")
sen <- read.table(fn_sen, header = FALSE, sep = "", dec = ".")
sen <- sen %>% rename(mon = V1, day = V2, yea = V3, hou = V4, mnt = V5, sec = V6, err = V7, sta = V8, bat = V9, ssp = V10, hed = V11, pit = V12, rol = V13, tmp = V14, a1 = V15, checksum = V16)

d <- 24*3600*as.numeric(as.Date(paste(sen$yea[1], sen$mon[1], sen$day[1], sep="-"), origin="1970-01-01")) # number of seconds that gives the day
h <- sen$sec[1]+60*sen$mnt[1]+3600*sen$hou[1] # time in seconds
starttime <- as_datetime(d + h) # lubridate datetime for the start of the data
fn_dat <- paste(fh, "dat", sep = ".")
dat <- read.table(fn_dat, header = FALSE, sep = "", dec = ".")
dat <- dat %>% rename(burst = V1, ensemble = V2, w = V3, u = V4, v = V5, amp1 = V6, amp2 = V7, amp3 = V8, snr1 = V9, snr2 = V10, snr3 = V11, corr1 = V12, corr2 = V13, corr3 = V14, p_dbar = V15, a1 = V16, a2 = V17, checksum = V18)
sampling_rate = 64
bar_s <- 15 # averaging window in seconds.  Should evaluate range from depth/mean representative velocity to entire period
bar <- round(bar_s * sampling_rate) # in indexed values [i], round() needed to ensure that it fits within the dataset
# gives decimal time for each data record
dat$time <- starttime + (c(0:(nrow(dat)-1)))/sampling_rate
atmos <- mean(1e4*dat$p_dbar[1:10]) # Pa, to subtract atmospheric pressure
dat$depth <- -(1e4*dat$p_dbar - atmos)/(9.81*997)

start <- as.numeric(ymd_hms("2021-05-28 16:57:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-05-28 16:59:00")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df528AR214 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p3 <- ggplot(df528AR214, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))
  
```

```{r warning=FALSE, echo=FALSE}
start <- as.numeric(ymd_hms("2021-05-28 17:03:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-05-28 17:05:00")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df528AR214 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p4 <- ggplot(df528AR214, aes(x= ui, y= wi)) +
  geom_point() +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))
```

```{r warning=FALSE, echo=FALSE}
start <- as.numeric(ymd_hms("2021-05-28 17:09:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-05-28 17:11:00")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df528AR214 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p5 <- ggplot(df528AR214, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))
```

```{r warning=FALSE, echo=FALSE}
start <- as.numeric(ymd_hms("2021-05-28 17:15:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-05-28 17:17:00")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df528AR214 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p6 <- ggplot(df528AR214, aes(x= ui, y= wi)) +
  geom_point() +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  xlim(-1, 1) +
  ylim(-1, 1) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))
```

```{r warning=FALSE, echo=FALSE}
grid.arrange(p1, p2, arrangeGrob(p3, p4, p5, p6, nrow=4, left = textGrob(TeX('$w_i$'), rot = 90, vjust = 1), bottom = textGrob(TeX('$u_i$'), vjust = 1)), nrow=1, top = "Allegheny River")
```

```{r warning=FALSE, echo=FALSE}
#Monongahela River
#cable crossing (CC)

#height vs DO
cable_crossing_DO <- read.csv("C:\\Users\\duquesne\\Documents\\Alanna\\DOcablecrossing629.csv")
cable_crossing_DO <- rename(cable_crossing_DO, time = ï..time) # This appears to be needed for Windows... 

p7 <- ggplot(cable_crossing_DO, aes(x = DO.mg.l. , y= height.m.))+
  geom_point() +
  labs(x = "DO (mg/l)", y = "Height(m)") +
  ggtitle("Dissolved Oxygen") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(axis.text = element_text(face = "plain", size = 8))
```

```{r warning=FALSE, echo=FALSE}
#height vs u
p8 <- ggplot(cable_crossing_DO, aes(x = U.m.s. , y= height.m.))+
  geom_point() +
  labs(x = "Average Velocity (m/s)", y = "Height(m)") +
  ggtitle("Velocity") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(axis.text = element_text(face = "plain", size = 8))
```

```{r warning=FALSE, echo=FALSE}
#stress tensors
#cable crossing goes from 15:09:00-15:23:00
setwd("c:/Users/duquesne/Documents/nortek/data")
fh <- "629MON15" # filename header
fn_sen <- paste(fh, "sen", sep = ".")
sen <- read.table(fn_sen, header = FALSE, sep = "", dec = ".")
sen <- sen %>% rename(mon = V1, day = V2, yea = V3, hou = V4, mnt = V5, sec = V6, err = V7, sta = V8, bat = V9, ssp = V10, hed = V11, pit = V12, rol = V13, tmp = V14, a1 = V15, checksum = V16)

d <- 24*3600*as.numeric(as.Date(paste(sen$yea[1], sen$mon[1], sen$day[1], sep="-"), origin="1970-01-01")) # number of seconds that gives the day
h <- sen$sec[1]+60*sen$mnt[1]+3600*sen$hou[1] # time in seconds
starttime <- as_datetime(d + h) # lubridate datetime for the start of the data
fn_dat <- paste(fh, "dat", sep = ".")
dat <- read.table(fn_dat, header = FALSE, sep = "", dec = ".")
dat <- dat %>% rename(burst = V1, ensemble = V2, w = V3, u = V4, v = V5, amp1 = V6, amp2 = V7, amp3 = V8, snr1 = V9, snr2 = V10, snr3 = V11, corr1 = V12, corr2 = V13, corr3 = V14, p_dbar = V15, a1 = V16, a2 = V17, checksum = V18)
sampling_rate = 64
bar_s <- 15 # averaging window in seconds.  Should evaluate range from depth/mean representative velocity to entire period
bar <- round(bar_s * sampling_rate) # in indexed values [i], round() needed to ensure that it fits within the dataset
# gives decimal time for each data record
dat$time <- starttime + (c(0:(nrow(dat)-1)))/sampling_rate
atmos <- mean(1e4*dat$p_dbar[1:10]) # Pa, to subtract atmospheric pressure
dat$depth <- -(1e4*dat$p_dbar - atmos)/(9.81*997)

start <- as.numeric(ymd_hms("2021-06-29 15:09:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-06-29 15:11:00")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df629MON15 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p9 <- ggplot(df629MON15, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))

```

```{r warning=FALSE, echo=FALSE}

start <- as.numeric(ymd_hms("2021-06-29 15:13:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-06-29 15:15:00")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df629MON15 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p10 <- ggplot(df629MON15, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))
```

```{r warning=FALSE, echo=FALSE}
start <- as.numeric(ymd_hms("2021-06-29 15:17:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-06-29 15:18:20")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df629MON15 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p11 <- ggplot(df629MON15, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-1, 1) +
  ylim(-1, 1) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))
```

```{r warning=FALSE, echo=FALSE}
start <- as.numeric(ymd_hms("2021-06-29 15:21:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-06-29 15:22:30")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df629MON15 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p12 <- ggplot(df629MON15, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))

```

```{r warning=FALSE, echo=FALSE}
grid.arrange(p7, p8, arrangeGrob(p9, p10, p11, p12, nrow=4, left = textGrob(TeX('$w_i$'), rot = 90, vjust = 1), bottom = textGrob(TeX('$u_i$'), vjust = 1)), nrow=1, top = "Monongahela River(cable crossing)")
```

   
```{r warning=FALSE, echo=FALSE}
#birmingham bridge (BB)
#height vs DO
birmingham_bridge_DO <- read.csv("C:\\Users\\duquesne\\Documents\\Alanna\\DOBirminghamBridge629.csv")
birmingham_bridge_DO <- rename(birmingham_bridge_DO, time =ï..time)

p13 <- ggplot(birmingham_bridge_DO, aes(x = DO.mg.l. , y= height.m.))+
  geom_point() +
  labs(x = "DO (mg/l)", y = "Height (m)") +
  ggtitle("Dissolved Oxygen") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(axis.text = element_text(face = "plain", size = 8))
```

```{r warning=FALSE, echo=FALSE}
#height vs u
p14 <- ggplot(birmingham_bridge_DO, aes(x = U.m.s. , y= height.m.))+
  geom_point() +
  labs(x = "Average Velocity(m/s)", y = "Height (m)") +
  ggtitle("Velocity") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(axis.text = element_text(face = "plain", size = 8))
```

```{r warning=FALSE, echo=FALSE}
#stress tensors

#birmingham bridge goes from 15:35:00-15:49:00
setwd("c:/Users/duquesne/Documents/nortek/data")
fh <- "629MON15" # filename header
fn_sen <- paste(fh, "sen", sep = ".")
sen <- read.table(fn_sen, header = FALSE, sep = "", dec = ".")
sen <- sen %>% rename(mon = V1, day = V2, yea = V3, hou = V4, mnt = V5, sec = V6, err = V7, sta = V8, bat = V9, ssp = V10, hed = V11, pit = V12, rol = V13, tmp = V14, a1 = V15, checksum = V16)

d <- 24*3600*as.numeric(as.Date(paste(sen$yea[1], sen$mon[1], sen$day[1], sep="-"), origin="1970-01-01")) # number of seconds that gives the day
h <- sen$sec[1]+60*sen$mnt[1]+3600*sen$hou[1] # time in seconds
starttime <- as_datetime(d + h) # lubridate datetime for the start of the data
fn_dat <- paste(fh, "dat", sep = ".")
dat <- read.table(fn_dat, header = FALSE, sep = "", dec = ".")
dat <- dat %>% rename(burst = V1, ensemble = V2, w = V3, u = V4, v = V5, amp1 = V6, amp2 = V7, amp3 = V8, snr1 = V9, snr2 = V10, snr3 = V11, corr1 = V12, corr2 = V13, corr3 = V14, p_dbar = V15, a1 = V16, a2 = V17, checksum = V18)
sampling_rate = 64
bar_s <- 15 # averaging window in seconds.  Should evaluate range from depth/mean representative velocity to entire period
bar <- round(bar_s * sampling_rate) # in indexed values [i], round() needed to ensure that it fits within the dataset
# gives decimal time for each data record
dat$time <- starttime + (c(0:(nrow(dat)-1)))/sampling_rate
atmos <- mean(1e4*dat$p_dbar[1:10]) # Pa, to subtract atmospheric pressure
dat$depth <- -(1e4*dat$p_dbar - atmos)/(9.81*997)

start <- as.numeric(ymd_hms("2021-06-29 15:35:40")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-06-29 15:37:00")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df629MON15 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p15 <- ggplot(df629MON15, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))

```

```{r warning=FALSE, echo=FALSE}
start <- as.numeric(ymd_hms("2021-06-29 15:39:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-06-29 15:41:00")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df629MON15 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p16 <- ggplot(df629MON15, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))

```

```{r warning=FALSE, echo=FALSE}
start <- as.numeric(ymd_hms("2021-06-29 15:43:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-06-29 15:45:00")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df629MON15 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p17 <- ggplot(df629MON15, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))
```

```{r warning=FALSE, echo=FALSE}
start <- as.numeric(ymd_hms("2021-06-29 15:47:00")) # Enter start time here as "YYYY-MM-DD HH:MM:SS" in 24-hour time
end <- as.numeric(ymd_hms("2021-06-29 15:48:30")) # End time, same format
for (i in 1:nrow(dat)) {
      if (as.numeric(dat$time[i]) < start) {
            s <- i
      }
      if (as.numeric(dat$time[i]) < end) {
            e <- i
      }
}
s <- s + 1
e <- e + 1
u_ave <- mean(dat$u[s:e])
v_ave <- mean(dat$v[s:e])
w_ave <- mean(dat$w[s:e])
ui <- dat$u[s:e] - u_ave
vi <- dat$v[s:e] - v_ave
wi <- dat$w[s:e] - w_ave
uiuj <- array(NA, dim = c(3,3))
uiuj[1,1] <- mean(ui^2)
uiuj[1,2] <- mean(ui*vi)
uiuj[1,3] <- mean(ui*wi)
uiuj[2,2] <- mean(vi^2)
uiuj[2,3] <- mean(vi*wi)
uiuj[3,3] <- mean(wi^2)

df629MON15 <- data.frame(ui, vi, wi)#creates data frame

#plots at each depth
p18 <- ggplot(df629MON15, aes(x= ui, y= wi)) +
  geom_point() +
  xlim(-0.5, 0.5) +
  ylim(-0.5, 0.5) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1)+
  theme(axis.text = element_text(face = "plain", size = 4.5))
```

```{r warning=FALSE, echo=FALSE}
grid.arrange(p13, p14, arrangeGrob(p15, p16, p17, p18, nrow=4, left = textGrob(TeX('$w_i$'), rot = 90, vjust = 1), bottom = textGrob(TeX('$u_i$'), vjust = 1)), nrow=1, top = "Monongahela River (Birmingham Bridge)")
```


    